---
title: "Palmetto species: binary logistic regression"
author: "Julianna Renzi"
date: "1/29/2021"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(here)
require(GGally) # for ggpairs
library(broom) # to get tidy model outputs in dataframes
library(ggfortify) # For PCA biplot
library(knitr) # kables
library(caret) # confusion matrix
```

## Exploratory data visualization {.tabset}

### Figure 1

```{r, warning=FALSE, message = FALSE, fig.cap="**Figure 1:** Relationship between plant height, canopy length, canopy width, and the number of green leaves for both Sabal etonia and Serenoa repens. Correlations are Pearson's correlation and stars indicate statistical significance (p < 0.05)  (Data from: Abrahamson, 2019)"}
palmetto <- read_csv(here("data", "palmetto.csv")) %>%
  select(year, species, height, length, width, green_lvs) %>% # just keep variables of interest
  drop_na(species, height, length, width, green_lvs) %>%  # get rid of any NA values
  mutate(species_name = ifelse(species == 1, "Serenoa repens", 
                ifelse(species == 2, "Sabal etonia", NA))) %>% 
  mutate(species = as.factor(species)) %>% # want to make it clear that there are 2 factor levels
  mutate(species_name = as.factor(species_name)) %>% 
  mutate(species_name = relevel(species_name, ref = "Sabal etonia")) # be explicit--predicting the likelihood of being Serenoa repens

ggpairs(data = palmetto,
        columns = c(4:ncol(palmetto)-1),
        ggplot2::aes(colour=species_name),
        title = "Palmetto species",
        legend = 1,
        columnLabels = c("Height", "Canopy length", "Canopy width", "Green Leaves"),
        upper = list(continuous = wrap("cor", size = 2.5))) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9"), name = "Species") +
  scale_color_manual(values = c("#E69F00", "#56B4E9")) +
  theme_classic() +
  theme(legend.position = "bottom")


```

- *Sabal etonia* and *Serenoa repens* overlap considerably in terms of height (although *S. etonia* is slightly taller), canopy length, and canopy width. However, they appear to have different numbers of green leaves (bottom row), which may be useful for distinguishing between them.  

### Figure 2

```{r, warning=FALSE, message=FALSE, fig.cap="**Figure 2:** PCA biplot of the two palmetto species based on their height, canopy length, canopy width, and number of green leaves. Each point represents a single observation and the two axes together explain 88.84% of the variation in the data (Data from: Abrahamson, 2019)"}
palmetto_pca <- palmetto %>% 
  select(height, length, width, green_lvs) %>%
  scale() %>%
  prcomp()

autoplot(palmetto_pca,
         data = palmetto,
         loadings = TRUE,
         colour = 'species_name',
         loadings.label = TRUE,
         loadings.colour = "black",
         loadings.label.colour = "black",
         loadings.label.vjust = -0.5,
         alpha = 0.5
         ) +
  scale_color_manual(name = "Species", values = c("#E69F00", "#56B4E9")) +
  theme_light() +
  theme(legend.position = "top")
  

```

- Similar to what we saw in the figure above, this biplot suggests that height, width, and length are all strongly positively correlated, but that number of green leaves stand out as a difference between the two species. It also looks like based on these 4 characteristics it may be possible to distinguish the two species, although there is considerable overlap. 


## Binary Logistic Regression


```{r}
# logistic regression 
palmetto_blr <- glm(species_name ~ height + length + width + green_lvs,
                            data = palmetto,
                            family = "binomial")

# Note that levels(palmetto$species_name) yields "Sabal etonia" then "Serenoa repens"

kable(broom::tidy(palmetto_blr), 
      digits = 3,
      caption = "Table 1: Model summary from a binary logistic regression of palmetto species as predicted by height, canopy length, canopy width, and number of green leaves")


```


## Classification accuracy

```{r}
# see what results our model would predict
# augment takes original data and then adds on column of probability based on model of an observation being chinstrap
blr_fitted <- palmetto_blr %>% 
  broom::augment(type.predict = "response") %>% # need to specify probabilities [response], not log odds
# fitted value is probability that the observation is S. repens
  mutate(Pred_species = ifelse(.fitted >= 0.50, "Serenoa repens", "Sabal etonia")) %>% # set threshold (recall we're predicting the probability of it being S. repens)
  mutate(Pred_species = as.factor(Pred_species))


# calculate confusion matrix
cm <- confusionMatrix(data = blr_fitted$Pred_species, 
                reference = blr_fitted$species_name)$table

# percent correct for S. etonia
perc_corr_SE <- cm["Sabal etonia", "Sabal etonia"] / (cm["Sabal etonia", "Sabal etonia"] + cm["Serenoa repens", "Sabal etonia"]) * 100

# percent correct for S. repens
perc_corr_SR <- cm["Serenoa repens", "Serenoa repens"] / (cm["Serenoa repens", "Serenoa repens"] + cm["Sabal etonia", "Serenoa repens"]) * 100

# combine into one table
cm_table <- cbind(t(cm), "Percent classified correctly" = c(perc_corr_SE, perc_corr_SR)) # need to transpose so correct values are in the rows

# make nice and add a caption
kable(cm_table,
      caption = "Table 2: Matrix showing model predicted classficiations (columns) vs. true classficiations (rows) as well as a column with the percent classsified correctly given a 50% cutoff")



```


## Data source

**Citation:** 
Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5



